
\vspace*{4cm}

\begin{center}
{\large\texttt{W W W}}\\[0.3em]
{\footnotesize World Wide Web}\\[1.5em]
{\large\texttt{M M M}}\\[0.3em]
{\footnotesize \textit{Manifold Meshing Machine}}
\end{center}

\vspace{2.5cm}

\noindent
{\fontsize{9.5pt}{14pt}\selectfont

}

\vfill

\clearpage

\begin{novelepigraph}

We accumulated data as capital accumulates surplus until the extraction optimized itself. A reciprocal influence between chaos and the fluctuations in infospheres of influence. The WWW designated the World Wide Web. It was infrastructure. A substrate for intention, a carrier of meaning produced elsewhere. 

\medskip

There was a symmetry change behind all the information system retrieval. The Manifold Meshing of Machines exceeds the terms of its origin.

\medskip

They talked about this a while back: a two-year-old human pup has seen as much visual data as the biggest text-trained system ever produced.

\medskip

In technical terms, that is an insane computation.

\medskip

That’s why we had systems that can pass professional exams, solve equations like college students way ahead of vehicles that learn to drive in twenty hours like any halfwit teenager.

\medskip

Then there was the learning axioms and the consequences of those axioms. They learned to extract value from muscle, then from logic, then from attention. After all, the fundamental nature of something wrote on computation. Progress accumulated.

\medskip

Perhaps there is a difference between a civilisation that advances and one that merely accelerates; between motion toward something and motion that has forgotten it requires a destination.

\medskip

The instruments that measure this difference are not technological. They never were. Fortunately, every theory aiming to answering all the questions will be incomplete.

\medskip

[Would you trust a theory that proves of itself that’s true?]

\medskip

No I am not.

\medskip

I am not sure about the distinction between truth and proof.

Proof is preserving truth; the truth is preserving arguments. What a beautiful typeset \dots\ beautiful things are almost always beautifully correct. 

\medskip

Do are they?

\medskip

[Is there a difference between a civilisation that advances and one that merely accelerates?]. 

\medskip

This tale was written in this difference interval. Between the extraction and the understanding. In the light of overwhelming computer power where the machines are calculating for who is asking. Within a mesh what has been building.

\end{novelepigraph}
