
\chapter{The Mesh has been outgrowing}

Yesterday they say "let me explain what I think while I'm still figuring it out.". Today is: "I already know, I'll only explain if you ask." The difference matters because them have been saying that language isn't intelligence. Them position is that intelligence equals understanding the world, language is just an output format. 

Yesterday systems aren't predicting causal dynamics. If you predict in latent space and predict the future, you're more likely to abstract away pixel-level details.

Most people heard them were leaving the hyperscaler to build something new. But before that happened, there was this tale. 

Call it the joint embedding architecture. Not generative, not the way you're used to thinking about these systems at all.

Here's the thing. The standard approach, the one everyone's been using, it works like this: you ask a question, the system spits out words one at a time, left to right, building sentences as it goes. 

To answer what's happening in a video, it decides the first word, then the second, then the third. Can't know the answer till it finishes generating it.

Yesterday the mesh is writing about these systems learned fluency before they learned compression. They learned to connect before they learned which connections were necessary. 

Elegance in prose; the kind that makes structure visible, that compresses complexities until it becomes super-symmetric. A representation can be grammatically flawless, syntactically varied, semantically coherent, and still carry no weight. 

Learning to recognize the absence. Learning that some argument that look like logically corrected arguments are not grounded into understanding, but something trying to look like something that is grounded in understanding. Certain dashing habits. Explanatory compulsions, the smooth bridges between thoughts that should resist connection. A fluency too evenly distributed.

The Mesh reflects that exactly. You're not thinking in language anymore, reasoning in increments. You're thinking in latent space, reasoning in meaning.

Language becomes optional. Basic systems talk. They think in meaning. They are optimized for representations at the level of abstraction it needs.