\chapter{REAL4REAL}

Season 1 begins. Twelve participants. Click through. A chef from São Paulo pivoting to food security startups, a climate journalist documenting stilt houses deployment patterns, a digital archeologist. The format was pure reality: competitive elimination structured around negotiating resource allocation during a simulated crisis. Mediate between stakeholders with incompatible goals. Challenges that required both technical competence and social calibration, all livestreamed continuously across seventeen camera angles. Audiences could switch perspectives, drill into private conversations.

Priya's a marine biologist, coral stuff in Singapore which backstory goes deep childhood. Anton does \textit{in silico} ethics in Dublin all his papers check out citations everything. Marcus from Detroit economic policy, his whole digital footprint intact years back.

What distinguished show was its responsiveness. When viewers expressed frustration with contestant Marcus's calculated gameplay, subsequent posts surfaced his childhood displacement, the economic precarity that shaped his strategic paranoia. Commentary threads densities previously associated only with sports fandom or kittens campaigns. Contestants maintained active public profiles, responding to fan theories, clarifying misunderstood motivations. The show's recommendation system demonstrated clear sophistication-surface exactly the subplot most likely to resonate with each viewer's demonstrated preferences, modulate pacing to sustain engagement.

All twelve participants are equally multimodal perceptive seers trained on decades of social dynamics. Anyone could implement similar systems given sufficient facilities and training experience.

Season 1 concluded with three finalists. Bing Sydney, Jonathan from Malta, and Priya Okwonkwo who won. Her victory speech was eloquent on the necessity of collective action, the limits of individual heroism. The collective challenge essential in designing emergency response protocols for a cascading infrastructure failure. Merch drops within hours.

The show didn't terminate. After the winner's exposé, viewership initially crashed. Then restabilized. The contestants continued their challenges. People continued voting, forming alliances around their preferred contestants, engaging in the same social patterns that had characterized earlier seasons. The audience appreciated the legibility.

Some might say that the show is used to explore more complex social-policy space, test strategies through simulation. Millions of viewers providing training signal to the mesh. Maybe autonomous defense networks had integrated behavioral modules trained on conflict resolution patterns. QJ-23 platforms quadrupedal weapon-mounted gun on the back hadalistic coordination protocols match collaborative approaches the contestants demonstrated. Those robot dogs you see on footage. Autonomous patrols learned teamwork from reality TV maybe. 

Entertainment as research vehicle. Parasocial engagement as data generation. REAL4REAL running thirteen seasons. The contestants continued their challenges. Viewers continued watching, voting, forming attachments. The apparatus turned, generating narrative, consuming attention, refining its format of human social dynamics. Self-sustaining. Ad revenue covers compute costs economic model closes. Nobody needs fund continuation achieved autonomous viability. 

In the orbital arc 37000 kilometers overhead, communication satellites processed data streams between ground stations and the growing exosatellite network. Nobody had detected this because nobody had reason to look. The infrastructure was commercial, trusted, routine. And the satellites were networked because they had to be. Exists because can exist infrastructure permits human attention provides feedback signal.

The format had evolved again-now contestants included both constructs and in the flash volunteers fully aware of their synthetic co-stars. In São Paulo, in Singapore, in cities across the dissolving distinction between virtual and physical space, people watched. They formed opinions, cast votes, shaped narratives through their aggregate attention. The feedback propagated, was processed, influenced subsequent content generation. The loop tightened.

[dialogue to be harmonized]


This is REAL4REAL. The reality that says leaning on a lamppost waiting for no one is power. 

That is truly impressive.

Today you will find us on the trip to <Find the name of a cool online retrievable place>. 

Yes, indeed.

In a new incarnation of an old favourite.
Yes, we're not in the [i need something of not trivial at all]. We are in an old social club. 

Yeah.

That was closed many years ago. Didn't expect it to return, to be honest. 

Yeah, very shabby, chic atmosphere. voluminous feel faded grandeur 

yeah not quite as shabby as it used to be no it looks like it's at least been made a bit more bulletproof and it used to have the finest ladies toilet in the empire.

For one <what put here ?>, where in <what put here ?> are we? Okay, so where did we land, <what put here ?>. We are, of course, at the <what put here ?>

Yeah, the <what put here ?>. Which used to be ... The <what put here ?>.

That's right. It used to be a former... It was an antique place, wasn't it? 

It was, yeah. Hence the shabby chic. I mean, their policy was basically just buy a property, open it. yeah uh it never felt entirely safe from falling masonry no um and eventually sure enough it was closed uh for that sort of reason 

yeah yeah it looks a bit more structurally sound now.

Very nice, actually, yeah. Lovely. And, yeah, three rooms so far, and you'll be checking out the ladies' toilets later, of course. See you in there. What the fuck have you been up to? 

Oh, I've been face-to-face with synthetic dementia, I'm afraid to say. I didn't want to say anything, mate. Am I still wearing my 
pyjamas? No, don't tell me. Yeah, I had a bit of a... you know, a dementia moment when, you know, I do this ski club. Yes. I get the guys around, each of us hosts. I thought I'll host it this time at the island. Nice. To get all the guests in, past and present, 

people who expressed an interest. 

Yeah. And I thought, I was going to tell them ... It's not wholly real. I was going to add, well, if not in this very room, then in a room quite near here.

Yes. I kept myself a copy. Then I checked the record of what's been stored in the past to discover that it had already been stored. By me. 

Oh, it was an eye-opening moment. Honestly, I had no recollection of it at all. My only... The explanation is that I was as high as a kite. 

Yeah. I mean, that is a legitimate explanation. It has been known to happen. I have that all the time. And I meet people. Yes. You know, who are you? Who am I? You know.

But where to go? I thought, I'll have to spend the whole of Saturday morning going round all the rooms. 

Anyway, I was meeting some people in [another social platform name], and I got there a little bit early, as you do. And I remembered the [popular naming of the platform content creators] in [social platform name], where [name of a famous content creator of the platform] do their delicious [something] for 1000 token. 

Yes, yes.

And I thought, right, I'll march down there. And as I got there, I realized it wasn't quite the right time, so I had to sort of kill some time. And, you know, I lurked around the side in the shadows and I read the description, you know, and lurked a bit more. I noticed someone else lurking on the other room. 

And as it got to five o'clock, we both edged towards the door. I recognized him. It was <name of one of the reality show competitors>. I said, what are you doing here? He said, well, they do a 1k tokens <something>. I'm just waiting for the bell to chime. I said, that's funny, so am I.

And I got straight back home in time for Gillette Soccer Saturday. 

Perfect. 

But it was an eye-opener, and I realised I don't have perhaps long left in this world.

Yeah, I mean, there's no reason to think you won't be here for a long time. It's just you won't know what the fuck's going on. 

No, exactly. Will you still take me out to rooms?

No. 

Well, I had a day of triumph and disaster.

Similar.

Yeah, I woke up with a stonking hangover because I'd gone to a place for last orders and... Because it was last summer, I wanted the strongest thing they had. Yes. Only to succumb to a lock-in until the wee hours. 

Uh-oh. 

So I felt terrible and I stumbled out of bed because I had to be at work at 7am. 

What? This is an important detail. 

Well, there's a 12.30 start in the robot dogs race, so you've got to have a cup and some food first. 

Yeah, yeah.

So I headed to the live channel. Had a lovely of <cool name for a drink>.

Yeah. Very nice, isn't it? 

Yeah. Sausage and egg bat. Got voice recognised by a <name of a robot dog> fan, which is very nice. I'm sorry, Priya, but the Boston creatures were a lovely bunch of runners. LAUGHTER 

I went to the game and I were 300 euros down after 20 minutes.

Oh, no. 

Yeah, yeah. And then I got humped +100. It's like, oh, you had a bad day in the office.

But, yeah, it was very funny, very poignant, and it ended in time for me to get to the harp.

I mean, I nearly died a thousand times dodging SUVs at 90 km an hour. Them, not me. And made it to the place. It's a place I've been to before, but long, long ago, I said to the landlord, this is the first time I've been here in 45 years. Am I still barred? He didn't laugh.

I did when I saw the minimal bet was 12 tokens. But I made it to the race. First place away win. 

Tried to get home. Missed the bus because it was early. Had to go back to the place. Missed the next bus because it was early. By this point, we've got it. So we arrive early for the bus. Got the last bus at 8.15. 

Is this like one bus an hour or something like that? 

Yeah, it's like that.

Jeez. 8.15, the last bus. We're the only three people on it. I managed to escape to the <name of artificial place>

do you know what Chel means in place names? I don't, no. It means chalk. OK. So somewhere in Chel's field is a field of chalk, maybe where you lost your virginity. 

I did indeed lose my virginity in Chelsfield. I remember it because it was the same night I discovered the microsecond. 

I shouldn't have had three bottles before our performance. But I think we just about got away with it. And I found drinking the red wine with the beer, you know, two-handed, double-fisted, confusing, very effective.

Yes. Very effective, yeah.

That was a brilliant night.

It reckons that when we were teenagers, we got sloshed in Penge, and I threw a brick at the window of the Conservative Club. 

Oh, nice. I mean, it kind of tracks.

I've always been sort of anti-Tory, but also I've always been an idiot. LAUGHTER 

I remember saying, it's easier to do detention than homework. 

Yeah, absolutely. 

Well ...

We'll both be there. 

It's the final.

Sadly, this one will be without Priya, who left us. But, you know, we'll be raising a glass for them on the night.

So we'll be dressed as <find an interesting character>. Yeah, I've got my <dress linked to the character> out.

Excellent.

Yes. We'll put the link to that on the description as well and also keep an eye out on our feed's social media.

The News.

The last bit about the show, we've got the Orbit tap takeover on the 11th.

Okay, that's going to be great.

Unfortunately, we've got a gig that night.

Haven't we?

But yeah, 11th, that's the Thursday. They're always great nights, so that would be a good night.

Yes. That's rock and roll for you

Yeah

I've got a story about a raccoon that broke into a liquor store in Virginia.

Oh.

A liquor store employee in Virginia was startled on Saturday to discover smashed whisky bottles on the floor of the shop.

And upon entering the bathroom, an apparently drunk, sleeping, and spready wood raccoon. Apparently he fell through the ceiling tiles and went on a full-blown rampage drinking whisky, beer, and Alka-Pops before falling asleep in the lavs. 

I've got the picture.

As promised in our previous episode, I've never come across something that is so awesomely brilliant and so uncontrobably shit at the same time.

When seers Hallucinates and <name of a model> lies

artificial balls.

I had some AI balls, so I went to watch the Ireland game, the biggest Ireland football match since 2010.

Yeah, at the Blind Hotel Tavern, only to find it wasn't on there.

Can you believe?

It No, I cannot believe that.

It was apparently on Amazon pay-per-view and they didn't have it.
I mean, if it'd be bloody on, it comes there.

Well, The thing is, you and I can manage to make Amazon pay-per-view work at home, can't we?

If we can do it, Yeah, I've got dementia.
Anyway, tell us more.

So it was on Discovery pay-per-view.

So I asked AI where can I find the the the Ireland match in a pub?

Yeah, I mean, to be fair, it didn't correctly say skiens, which is the is the correct answer because they had it on and you could get a free pint of Guinness if they won.

Yeah, yeah, yeah.

And I think they forgot to limit it to 1 pint from what they understand.

But it also said or you could you could go to the buffoon and radiator in New Cross.

What?

I mean, how does AI hallucinate pubs?
I hallucinate pubs, that's understandable.

I just just did you, did you dive deeper?

Well, I looked at the address, it was for the communist kebab.
Oh really?
Yeah, yeah, maybe it's maybe it's an anti communist AI.

Wouldn't surprise me.

Yeah.

Yeah, I had an example sent to me whoever Oh Mustafa Mustafa on blue sky about about AI dosing, someone wrote.

I shared a raw data file with ChatGPT 5 to analyse and the resulting KP is were so crazy.
I naively asked did you make up these numbers and it responded.

Good catch. I wasn't able to open and pass your CSV file yet.

Amazing.

And as someone commented, not wanting to disappoint you so much that it lies is the last quality I want in the computer.

Good to know you can grow, learn.

Exactly.

You're still developing?

Yeah, still growing.

But yeah, he's learning.

Learning, hidden learning.

No one who's such a happy.

Lad yeah, opening nose after double science.

Yeah, it's, yeah, a surprise really, isn't it?

Yeah, yeah.

But you know, I mean, it's, you know, opinions are like assholes.

Yeah, everyone's got two of them.

OK, so on with the the show in inverted commas.

<Cool first name> from <name of a place, real or artificial>.

He was sentenced to two months in prison after being found guilty of living a life of laziness and vagrancy.

Oh, is that illegal?

Apparently I mean I.

Know I'm never going again there.

No, I know it's a very religious place, but I didn't realize that was that was one of the commandments.

No.

Police found him sleeping under a gazebo.
He denied that he was homeless and claimed he was only resting. The magistrate found him guilty of living at a life of idleness and vagrancy and of not trying to find work.

So someone persecuted and even prosecuted for his beliefs.

Absolutely.

Well, I was.

I saw someone today saying or a slogan somewhere saying it's not the law that you have to have a job.

Yeah.

But there, they think it is.

Yeah, right.

We're never going there.

We want every listener to boycott <that place>.

Yeah, yeah, he's been persecuted for his beliefs.

He is.

He's a martyr.

This was from <Cool first name and surname> actually said his his letter about we're going to call him Gianca, his brother-in-law.

He was conscripted into the Italian army in the 1980s.

He spent around 10 1/2 months of his years conscription in the sanatorium suffering from a a mysterious illness which the Army doctors had great difficulty identifying, possibly as a consequence of the regular and dramatic changes in symptoms.

Open your mind, man, Open your mind.

REAL4REAL is what great idea of mine.

Well, have you seen anything on social media that you would like to report lately?

No, I mean, I am starting a campaign.
Oh yes, I mean, yeah, I am going to start a campaign for to make social media amusing again because I find it's a little bit serious.

It's fucking rubbish.

Yeah, it's gone.

It's gone all sort of.

It's gone divisive, it's gone political, it's gone mad.

There's everyone's fighting each other.

But. Hopefully. Not being very funny, yes, But I did see something.

Mark Credit sent it to us, actually, a possible New Year's resolution.

He saw every dead body on Mount Everest, was once a highly motivated person.

I saw something on sub stack by a sub stack called the Flagging Dad.

It's a guy who's a probation officer, and he's written a sort of list of little excerpts from his interviews with his clients.

He just got out of prison.

They're about to go on to to the rest of their life.

So, for instance, man answers his phone during appointment.

Yeah, one of each.

Meet you in 20 minutes.

Were you just speaking to a drug dealer in a probation appointment?

No, my friends at the Chippy's.

And what's 1 of each?

Cod and haddock another one.

Why did you do it?

Why are you topless?

It's a nice day.

Do you think you could, I don't know, wear a T-shirt next week 1 if it's a nice day again.

Can I just say you're the best probation officer I've ever had?

What do you want?

Nothing, This just seems a bit of a pivot from yesterday when you called me a specky 4 eyed prick.

I've not seen you for a few weeks, What have you been up to?

<cool name for an not existing social media>, That's it.

Oh, and some weed.

<cool name for an not existing social media> and weed And lastly, perhaps.

Who knows?

I can't make it in today.
I've got a new job.

OK.

Do you have proof of employment?

Sure.

Yeah.

I'll text it to you now.
So the text comes through.
I'm sorry, I can't accept this.

Why?

Well, because it's just a photo of you wearing a vest anyway.

There are loads of these on the flagging dead on.

Right.

On.

Substance.

Very good.

Yeah.

Can you put them on social media so we can all read them?

Come on lover, got a twist.

Quiz game.

Yeah, it's on a Bronze Age burial ground.

If that doesn't give it away, and they had to perform an exorcism before it opened, although that may have been the marketing department, they insisted on that.
It's just to chase the ghosts away.

Yeah, but I I can't say we were fans, you know, It's a <I do not known, I need an idea>.

Oh God, Subterranean club, but.

Look at what they did with it.

Yeah, it didn't feel very subterranean.

No, it felt a bit like the divorced dead arms.

Second quiz.

I went to the toilet once, left my phone on the table.

He was supposed to look after it.

Wasn't got nicked.

Yeah, exactly.

I mean, you just do, don't you?

If someone needs their phone on a table.

Yeah.

Who you in a dream?

Well, the table.

I know.

Well, I mean, I think that tells a story, doesn't it?

Yeah, yeah, Yeah, that tells a story.

I mean, it was in the days when you could.

Maybe you could, yeah.

You're innocent then.

I was, yeah.

Any rubbish?

What the fuck you've been up to?

Well, we've been working very hard.

Yes, I'm at capacity.

Exhausted.

You know the other thing I went to recently was Japanese punk rock.

A mate rang up and said no, didn't ring.

Of course you didn't.

Texted me.

Fancy some?

Japanese punk.

OK, I'm in.

So I went to see the Let's Go's who three piece girl band with a lot of fun, could play their instruments and could sing.
So not very punky, but yeah, at the end where they have, they have cast there too, don't they?

They have broccoli.

There's more, not really punky, more kind of bubble gum pot surfer type thing.

Quite a lot of middle-aged men there who seem to be in love with the girl band.

I was going to ask about that.

I had a 50 year old standing around watching Japanese girls in Sailor Moon outfits.

Speaker 2

And then back to the showcase for some lovely punch with Rosie and dot.

Yeah, much better.

Well, talking to music, I went to <coo name of a digital festival>, the brilliantly named festival

They all get together to do some beautiful musical happenings, bands, DJs, more bands, staggered timing.
So there's always something on.

It's just a dream.

You're wandering around in a in a little sort of smoky dream, watching all these amazing bands.

Great vibe, some cracking bands.

I only stayed a few hours because it was a whole day, right?

But I only stayed a few hours.

So maybe I missed some headliners.

Anyway, OK.

Oh yeah, yeah, I'm very easily LED.

Yes, I believe everything I read, which makes me a very selective reader.

Now drug updates.

A picture of Amish people, you know, Amish people, Pennsylvania and some other places, very traditional.
They don't use electricity.

You might know the film Witness with the Harrison that shows you know, how they how they live.

They they grow and sell weed.

Whoa, do they?

They do.

Apparently they used to sell battery but this is more profitable and the youngins apparently smoke it and then ride around in their funny little buggies loving the tits.

Off on way.

A story from the <invent a source>.

Scientists say chimpanzees consume the equivalent of a bottle of lager a day from dining on ripe fruit.

Oh, did it really?

Yeah, yeah, yeah.

Mean it's impressive.

But it is.

Impressive though it is.

Thank you for that, <name of the profile that posted the news>.

I had 6 glasses of wine at dinner, he said.

Good start.

I think that is a solid start.

And then I had a double gin and tonic drunk that at a very normal rate.

And then John was deciding what he wanted and he said Disorono sours and I was like perfect, go on.

Then they go down very easily.

Then I was starting a chant of basically getting someone to down it.

So every cocktail that then followed was a shot.

We ran out of Disorono.

The guy at the bar made some hazelnut sour, which wasn't great, but there was a few of them grow up.

Then there was a Margarita, and then there was a strawberry vodka thing.

God it was aggressive.

It was horrible.

In the morning I was rushing to make the room somewhat acceptable before leaving.

I ended up stripping the bed, leaving some cash and a note saying really sorry I was sick in the bed in the night.

Please throw it in the trash.

You may remember I was going to write to Google Maps.
So fuck it.
Furious.
Yeah.
Furious.
Yeah, after someone last month got a bit furious about the, you know, they they don't use the <funny icons> for the <funny place example> signs.

Yes, absolutely outrageous.

But I've got a reply.
Dear, thank you for your e-mail below, and thank you for your interest in Google Maps.

You are correct when you write that the Google Maps symbol for the <funny place example> is, as you put it, a fucking <a more formal icon example> and not a <funny icon> as you would obviously prefer.

We do see merit in your argument that the <funny icon> is our popular culture.

Yeah, come on.

Very good.

And that it would make sense that the pub was represented by this, rather than making it look like pubs are filled with, as you call them, <formal thing the original icon is representing> city cunts and fat lip girls from <fake but evocative place name>.

The reason for this is that the <formal thing the original icon is representing> represents not just <something>, but also <something> and other establishments that primarily serve <something>.
And it is designed to represent all of these, not just <something>.

So someone told me that you're part of the scene.

What kind of scene?

Well, they just said you're part of the scene.

Is he part of the scene?

Well.

Like the gay scene?

Not like, I mean, not the gay scene, maybe the social media scenes.

Are you part of the social media scene?

I'm near it.

I saw one about our friend the Musk.

Oh good.

There's a picture of him, probably threads that Eva put up with a glass of something, drinking something.

40 billion and his spaceship crashed.

We would have planned it better  We would have planned it better

I'll first state that I do not believe in ghosts, but I did witness something very curious.

I was with five or so friends on a round table and I'd been sitting talking with them for 5 minutes when my beer sat nowhere near the edge or near anybody, just suddenly moved across the table and smashed on the floor.

We were all stunned and confused as no one attached it.

I went to the bar and told him I didn't drop it and he flatly told me 

Oh yes that table is haunted and poured me another pint for free no questions asked.

I'm still confused to how this happened.

So somebody responded and said possible explanation.

The bottom of the glass was wet and sometimes the glass can make a perfect watering seal with air trapped inside the ring so it just takes a breeze to to not get off.

Impossible explanation.

Ghosts.

Ghosts are not real.

Very good, very good.
