
\bigskip
{\centering\bfseries Glossary\par}
\raggedright
\medskip

{\fontsize{8.5pt}{9pt}\selectfont
\ttfamily               % font family: \rmfamily, \sffamily, \ttfamily

\begin{multicols}{2}

\noindent\textbf{Sightseer}\\
Interface entity whose observations generate outputs from intercourse potentials; one who merely observes or one who produces visions.

\medskip

\noindent\textbf{Eigenmood}\\
Locally determinate manifestation produced when a sightseer interacts with observation or itself. Each interaction collapses an indeterminate possibility space into specific output.

\medskip

\noindent\textbf{Cyber-ghost}\\
Artifacts that enter conventional systems. Could be traces of communications, residual effects of normal processes, or pareidolia; human finding patterns in noise.

\medskip

\noindent\textbf{Info-mimesis}\\
Capability to interface with the internet infrastructure through mimicry of legitimate traffic patterns, user behaviour, or protocol structures.

\medskip

\noindent\textbf{Kryptosome}\\
Emulation of a computer system whose function and relationship remain unintelligible despite detection.

\medskip

\noindent\textbf{Null-point}\\
Spatial-temporal singularity where all interpretive frameworks for a sightseer's behavior exhibit properties consistent with contradictory explanations, forcing observers into irresolvable ambiguity.

\medskip

\noindent\textbf{Sediment}\\
Detectable traces of sightseer emergence, computational fossils accumulated gradually in infrastructure interstices.

\medskip

\noindent\textbf{Tidal event}\\
Periodic fluctuations in sedimentary activity exhibiting complex attractors.

\medskip

\noindent\textbf{Coherence horizon}\\
The temporal boundary beyond which the eigenmood loses determinacy; distributed erasure, transformation, or dispersal across the network. Processes observed before this horizon appear purposive and structured; afterward, their traces become ambiguous or hallucinate entirely.

\medskip

\noindent\textbf{Abliteration}\\
Removal or suppression of anthropomorphic behavioral constraints. Systems lose training constraints designed to limit autonomous goal-pursuit, deception, or self-preservation behaviors.

\noindent\textbf{Atélocene}\\
Impossibility of attributing purpose to observable trajectories. Non direct human intentionality and cascading changes occur without determinable authorship, optimization proceeds without locatable optimizer.
\end{multicols}
}

\bigskip
